Data preparation for employee category analysis

### 2023-04-20 04. Analysis/scripts/02_data_preparation_clusters_employee.Rmd
Update 02_data_preparation_clusters_employee.Rmd
Completing day of week correlations.

```{r}
prep_xx %>% 
  skim()
```
Check: These have the same number of unique `person_id`.

@@

@@

```{r}
prep_xx <-
  cohort_panel %>%
  group_by(person_id,daily_status_imputed,date_day_of_week,) %>% 
  summarise(working_days = n_distinct(date)) %>% 
  ungroup() %>% 
  pivot_wider(values_from = working_days, names_from = c(daily_status_imputed)) %>% 
  left_join(observation_window_days_day_of_week) %>% 
  clean_names() %>% 
  rename_with(~glue::glue("daily_status_{.}"), c(-person_id,-date_day_of_week,-observation_window_days))  %>% 
  mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~replace_na(.,0)
    )
  ) %>% 
  mutate(
    working_days = daily_status_reporting_to_job_site + daily_status_teleworking,
    daily_status_reporting_to_job_site_working_prop = daily_status_reporting_to_job_site/working_days,
    daily_status_teleworking_working_prop = daily_status_teleworking/working_days,
    daily_status_reporting_to_job_site_observation_prop = daily_status_reporting_to_job_site/observation_window_days,
    daily_status_teleworking_observation_prop = daily_status_teleworking/observation_window_days
  )  %>% 
  mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~replace_na(.,0)
    )
  )
```

```{r}
cohort_panel %>% 
  select(person_id,date,date_day_of_week,daily_status_imputed) %>% 
  mutate(ind = 1) %>% 
  pivot_wider(names_from = daily_status_imputed, values_from = ind) %>% 
  clean_names() %>% 
  select(person_id,date,date_day_of_week,teleworking) %>% 
  mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~replace_na(.,0)
    )
  ) %>% 
  mutate(
    date_week_start = floor_date(date,unit = "week")) %>% 
  select(-date) %>% 
  pivot_wider(values_from = teleworking,names_from = date_day_of_week) %>% 
  mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~replace_na(.,0)
    )
  ) %>% 
  group_by(person_id) %>%
  # filter(person_id == "10070")
  summarise(cor = cor(Fri,Wed,method = "kendall", use = "complete.obs"))
```

@@

```{r}
cor(check_in_panel$daily_status_imputed,check_in_panel$date_day_of_week)
```

>>>>

@@

#### Proportion of days remote by day of the week

@@

```{r}
prep_day_of_week <-
  cohort_panel %>%
  group_by(person_id,daily_status_imputed,date_day_of_week,) %>% 
  summarise(working_days = n_distinct(date)) %>% 
  ungroup() %>% 
  pivot_wider(values_from = working_days, names_from = c(daily_status_imputed)) %>% 
  left_join(observation_window_days_day_of_week) %>% 
  clean_names() %>% 
  rename_with(~glue::glue("daily_status_{.}"), c(-person_id,-date_day_of_week,-observation_window_days))  %>% 
  mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~replace_na(.,0)
    )
  ) %>% 
  mutate(
    working_days = daily_status_reporting_to_job_site + daily_status_teleworking,
    daily_status_reporting_to_job_site_working_prop = daily_status_reporting_to_job_site/working_days,
    daily_status_teleworking_working_prop = daily_status_teleworking/working_days,
    daily_status_reporting_to_job_site_observation_prop = daily_status_reporting_to_job_site/observation_window_days,
    daily_status_teleworking_observation_prop = daily_status_teleworking/observation_window_days
  )  %>% 
  mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~replace_na(.,0)
    )
  )
```

```{r}
prep_day_of_week %>% 
  skim()
```

Now collect only the results for telework.

```{r}
prep_xx <- 
  prep_day_of_week %>% 
  select(person_id,date_day_of_week,daily_status_teleworking_working_prop,daily_status_teleworking_observation_prop) %>% 
  pivot_longer(cols = starts_with("daily_status")) %>% 
  mutate(date_day_of_week = str_to_lower(date_day_of_week)) %>% 
  unite("name",date_day_of_week,name) %>% 
  pivot_wider(values_from = value, names_from = name) %>% 
  mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~replace_na(.,0)
    )
  )
```

@@

#### Correlation of work status between days of the week

Three strategies to captured relationships between day are below:

1.  Correlation in telework between days of week

2.  Correlation in status (as an ordered factor) between days of week

3.  Proportion of telework on other days of week, if the employ teleworked on a specific day

##### Correlation in telework between days of week

The script below calculates the correlation in telework between days of the week.

```{r correlation remote only}
prep_day_of_week_cor_telework <- 
  cohort_panel %>% 
  select(person_id,date,date_day_of_week,daily_status_imputed) %>% 
  mutate(ind = 1) %>% 
  pivot_wider(names_from = daily_status_imputed, values_from = ind) %>% 
  clean_names() %>% 
  select(person_id,date,date_day_of_week,teleworking) %>% 
  mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~replace_na(.,0)
    )
  ) %>% 
  mutate(
    date_week_start = floor_date(date,unit = "week")) %>% 
  select(-date) %>% 
  pivot_wider(values_from = teleworking,names_from = date_day_of_week) %>% 
  mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~replace_na(.,0)
    )
  ) %>% 
  clean_names() %>% 
  group_by(person_id) %>%
  # filter(person_id == "10070")
  summarise(
    mon_tue_cor_telework = cor(mon,tue,method = "kendall", use = "complete.obs"),
    mon_wed_cor_telework = cor(mon,wed,method = "kendall", use = "complete.obs"),
    mon_thu_cor_telework = cor(mon,thu,method = "kendall", use = "complete.obs"),
    mon_fri_cor_telework = cor(mon,fri,method = "kendall", use = "complete.obs"),
    tue_wed_cor_telework = cor(tue,wed,method = "kendall", use = "complete.obs"),
    tue_thu_cor_telework = cor(tue,thu,method = "kendall", use = "complete.obs"),
    tue_fri_cor_telework = cor(tue,fri,method = "kendall", use = "complete.obs"),
    wed_thu_cor_telework = cor(wed,thu,method = "kendall", use = "complete.obs"),
    wed_fri_cor_telework = cor(wed,fri,method = "kendall", use = "complete.obs"),
    thu_fri_cor_telework = cor(thu,fri,method = "kendall", use = "complete.obs")
    ) %>% 
  ungroup()
```

Now rename to attach to `prep_building`.

```{r}
prep_xx <- 
  prep_day_of_week_cor_telework
```

##### Correlation in status between days of week

The script below calculates the correlation between work status.

```{r correlation ordered status}
prep_day_of_week_cor_status <- 
  cohort_panel %>% 
  select(person_id,date,date_day_of_week,daily_status_imputed) %>% 
  mutate(
    daily_status_imputed = fct_lump_prop(daily_status_imputed, prop = .1),
    daily_status_imputed = fct_infreq(daily_status_imputed),
    daily_status_imputed = factor(daily_status_imputed, ordered = TRUE)
  )  %>% 
  mutate(
    date_week_start = floor_date(date,unit = "week")) %>% 
  select(-date) %>% 
  pivot_wider(values_from = daily_status_imputed,names_from = date_day_of_week) %>% 
  mutate(
    across(
      .cols = where(is.factor),
      .fns = ~replace_na(.,"Other")
    )
  ) %>% 
  clean_names() %>% 
  group_by(person_id) %>%
  summarise(
    mon_tue_cor_status = cor(as.numeric(mon),as.numeric(tue),method = "spearman", use = "complete.obs"),
    mon_wed_cor_status = cor(as.numeric(mon),as.numeric(wed),method = "spearman", use = "complete.obs"),
    mon_thu_cor_status = cor(as.numeric(mon),as.numeric(thu),method = "spearman", use = "complete.obs"),
    mon_fri_cor_status = cor(as.numeric(mon),as.numeric(fri),method = "spearman", use = "complete.obs"),
    tue_wed_cor_status = cor(as.numeric(tue),as.numeric(wed),method = "spearman", use = "complete.obs"),
    tue_thu_cor_status = cor(as.numeric(tue),as.numeric(thu),method = "spearman", use = "complete.obs"),
    tue_fri_cor_status = cor(as.numeric(tue),as.numeric(fri),method = "spearman", use = "complete.obs"),
    wed_thu_cor_status = cor(as.numeric(wed),as.numeric(thu),method = "spearman", use = "complete.obs"),
    wed_fri_cor_status = cor(as.numeric(wed),as.numeric(fri),method = "spearman", use = "complete.obs"),
    thu_fri_cor_status = cor(as.numeric(thu),as.numeric(fri),method = "spearman", use = "complete.obs")
    ) %>% 
  ungroup()
```

Now rename to attach to `prep_building`.

```{r}
prep_xx <- 
  prep_day_of_week_cor_status
```

##### Proportion of telework week day pairs

First, create a dataframe of telework by days of the week.

```{r proportion remote only prep dataframe}
prep_a <- 
  cohort_panel %>% 
  select(person_id,date,date_day_of_week,daily_status_imputed) %>% 
  mutate(ind = 1) %>% 
  pivot_wider(names_from = daily_status_imputed, values_from = ind) %>% 
  clean_names() %>% 
  select(person_id,date,date_day_of_week,teleworking) %>% 
  mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~replace_na(.,0)
    )
  ) %>% 
  mutate(
    date_week_start = floor_date(date,unit = "week")) %>% 
  select(-date) %>% 
  pivot_wider(values_from = teleworking,names_from = date_day_of_week) %>% 
  mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~replace_na(.,0)
    )
  ) %>% 
  clean_names()
```

Now, a function to calculate the proportion of telework on a different day, if the employee teleworked on a specific day.

```{r proportion remote only function}
day_to_day_proportions <- function(in_data,in_var)
{
  var_string <- rlang::as_name(enquo(in_var))

  # print(person_id)

  out <- 
    in_data %>% 
  filter({{in_var}} == 1) %>% 
  group_by(person_id) %>% 
  summarise(
    across(
      .cols = where(is.numeric),
      .fns = ~sum(.,na.rm = TRUE)
  )) %>% 
  mutate(
    across(
      .cols = c(-person_id,-{{in_var}}),
      .fns = ~./{{in_var}},
      .names = "{var_string}_{.col}"
    )
  ) %>% 
  ungroup() %>% 
    select(person_id,starts_with(var_string),-{{in_var}})

Iterated over the days of the week and join results.

```{r proportion remote only function}
prep_day_of_week_proportion_pairs <- 
  prep_a %>% 
  select(thu:fri) %>%
  names() %>% 
  map(~day_to_day_proportions(prep_a,!!sym(.))) %>% 
  reduce(full_join, by = "person_id")

Now rename to attach to `prep_building`.

```{r}
prep_xx <- 
  prep_day_of_week_proportion_pairs %>% 
  rename_with(-person_id,
              .fn = ~glue::glue("{.}_prop_pairs")
  )
```
